---
title: "Untitled"
format: html
editor: visual
---

### Libraries

```{r}
library(bayesrules)
library(tidyverse)
library(ggpubr)
library(plotly)
library(report)
library(ggplot2)

```

### Intro

Welcome to today's lecture, we will analyze Capital Bikeshare Data from Washington D.C

```{r}
bikes_data <- bayesrules::bikes
```

Starter Questions:

1.  What is the distribution of ridership? Make a histogram and write what you see below

```{r}
ggplot(data = bikes_data, aes(x = rides)) +
  geom_histogram(color = 'black') +
  theme_bw()
```

2.  Has ridership increased over time? Make a plot and write what you see below

```{r}
ggplot(data = bikes_data, aes(x = date, y = rides)) +
  geom_line() 
```

```         
```

```{r}
ggplot(data = bikes_data, aes(x = rides, fill = month)) +
  geom_histogram(color = 'black') +
  theme_bw()
```

```{r}
line_plot <- ggplot(data = bikes_data, aes(x = date, y = rides, label = day_of_week)) +
  geom_line() +
  geom_point() 
ggplotly(line_plot)
```

```{r}
ggplot(bikes_data, aes(x = as.factor(year), fill = as.factor(year), y = rides)) +
    geom_boxplot() +
    geom_jitter()
```

3.  Are there more rides on weekdays or weekends? Make a plot and write what you see below

```{r}
ggplot(bikes_data, aes(x = weekend, fill = weekend, y = rides)) +
  geom_boxplot() +
  geom_jitter()
```

```{r}
ggplot(bikes_data, aes(x = weekend, fill = weekend, y = rides)) +
  geom_col() 
```

4.  **Does temperature feel effect ridership**?

```{r}
ggplot(bikes_data, aes(x = temp_feel, y = rides)) +
  geom_point()
```

```{r}
ggplot(bikes_data, aes(x = temp_feel, y = rides)) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_cor()
  
```

What is a correlation coefficient and p value: [helpful tool here](https://rpsychologist.com/correlation/)

R = 0.58 — the strength and direction of the pattern you're looking at

- On the -1-to-1 scale, 0.58 sits in **moderate** territory: real and clearly visible in the scatter, but the cloud of points still has meaningful width around the line, not tightly hugging it the way an R of 0.9+ would.

Squaring it tells you something more concrete about the plot: R² ≈ 0.34

- About **34% of the vertical scatter in ridership** across this plot is explained by where a point sits on the x-axis (temperature). The remaining **66%** is everything else the plot doesn't capture

p \< 2.2×10⁻¹⁶ — how confident you should be that this pattern is real, not a fluke of these particular 500 days

- This tests rhe hypothesis that temperature and ridership have *no* linear relationship at all in the underlying population, and this pattern is just coincidental noise from sampling 500 particular days. The p-value says: **if that null hypothesis were actually true, seeing a correlation as strong as 0.58 in a sample this size would be astronomically unlikely**

```{r}
lm(data = bikes_data, rides~temp_feel)
```

**`(Intercept)` = -2179.27**: the model's predicted ridership when temp_feel = 0. As flagged before, this is a real number the formula produces, but not a practically meaningful one — 0°F essentially never happens in D.C., so this is just "where the line would cross the y-axis if you extended it that far," not a usable prediction.

**`temp_feel` = 81.88**: the slope — for every one-unit (one degree) increase in temp_feel, the model predicts ridership increases by 81.88 rides. This is the number that actually answers "does temperature matter, and by how much"

```{r}

model <- lm(rides ~ temp_feel, data = bikes_data)
report(model)
```

### Your turn

Repeat the all the steps to model the relationship between

- rides and humidity

```{r}
ggplot(data = bikes_data, aes(x = humidity, y = rides)) +
  geom_point(color = 'lightgreen') +
  geom_smooth(method = 'lm') +
  stat_cor()
```

```{r}
lm(data = bikes_data, rides~humidity)
```

```{r}

```

- rides and windspeed

```{r}
ggplot(data = bikes_data, aes(x = windspeed, y = rides)) +
  geom_point(color = 'lightblue' ) +
  geom_smooth(method = 'lm') +
  stat_cor()
```

```{r}
lm(data = bikes_data, rides~windspeed)
```

- rides and temp_actual

```{r}
ggplot(data = bikes_data, aes(x = temp_actual, y = rides)) +
  geom_point(color = 'lightpink') +
  geom_smooth(method = 'lm') +
  stat_cor()
```

```{r}
lm(data = bikes_data, rides~temp_actual)
```

- temp_feel and temp_actual

```{r}
ggplot(data = bikes_data, aes(x = temp_actual, y = temp_feel)) +
  geom_point(color = 'black') +
  geom_smooth(method = 'lm') +
  stat_cor()
```

```{r}
lm(data = bikes_data, temp_feel~temp_actual)
```
